---
title: "Evaluating and Projecting Progress with trackr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Evaluating and Projecting Progress with trackr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

```{r setup}
library(trackr)
```

## Getting Started

Installation

You can install trackr directly from GitHub:

```{r}
# install.packages("devtools")
#devtools::install_github("RossanaTat/trackr")

```

Then load it:

```{r}
library(trackr)

```

Input Data Structure:

`track_progress()` expects a dataset with at least three columns:
`code_col`, `year_col` and `indicator` Overview

The trackr package helps users evaluate and project progress on any
indicator across countries and time. It provides two complementary
scoring approaches:

Speed of progress: how fast a country improved relative to a typical
experience

Percentile score: how a country’s progress compares to the distribution
of all observed experiences

Together, these methods allow analysts to assess historical trends and
simulate future trajectories under different assumptions.

## 🌍 Example: Tracking Undernourishment (WDI)

In this example, we demonstrate how **trackr** can be used with real
data from the [World Development Indicators
(WDI)](https://databank.worldbank.org/source/world-development-indicators)
to evaluate and project progress on a global development indicator.

We’ll use the indicator **“Prevalence of undernourishment (% of
population)”**,\
identified in WDI as `SN.ITK.DEFC.ZS`. Lower values indicate better
outcomes.

```{r}
# Example 1: Tracking progress on undernourishment (% of population)
indicator_wdi <- "SN.ITK.DEFC.ZS"

data_wdi <- wbstats::wb_data(
  indicator = indicator_wdi,
  lang      = "en",
  country   = "countries_only"
)

results_progress <- suppressWarnings(track_progress(
                        data           = data_wdi,
                        indicator      = indicator_wdi,
                        code_col       = "iso3c",
                        year_col       = "date",
                        startyear_data = 1960,
                        endyear_data   = 2024,
                        eval_from      = 2015,
                        eval_to        = 2024,
                        future         = TRUE,
                        target_year    = 2030,
                        speed          = TRUE,
                        percentiles    = TRUE,
                        sequence_pctl  = seq(20, 80, 20),
                        sequence_speed = c(0.25, 0.5, 1, 2, 4),
                        best           = "low",  # lower undernourishment = better
                        min            = 0,
                        max            = NULL,
                        support        = 1,
                        granularity    = 0.1
                      ))

```

### 📈 Interpreting the Output

The output is a list of up to six components, depending on the user’s
settings. Here we’ve enabled both speed and percentile methods, as well
as future projections.

```{r}
names(results_progress)
```

Examples of relevant results for interpretation are:

```{r}
results_progress$scores # the country-level progress scores for 2015–2024

results_progress$path_future # projections to 2030 under different assumptions

head(results_progress$scores)
```

### 🔍 Understanding the Scores

A speed score greater than 1 means a country reduced undernourishment
faster than the typical global trend. For instance, 2.0 implies progress
twice as fast as average.

A percentile score of 80 means the country outperformed 80% of
comparable experiences.

Together, these metrics help identify which countries are making
accelerated progress and which may need additional support to meet
global nutrition goals

🔮 Visualizing XXXX TODO: Do we want to add a little vis here?

This plot shows ... FEW WORDS HERE

## Key Takeaways

trackr can work directly with WDI or any other dataset structured with
country-year-indicator values.

It produces transparent, reproducible progress assessments using
comparable methods across indicators and countries.

These outputs are particularly useful for monitoring development goals,
benchmarking progress, and exploring future target scenarios.

## Summary

This example illustrated how to use track_progress() with WDI data to
assess and project progress on undernourishment. The same workflow can
be applied to any other indicator — either simply changing the WDI code
or providing your own data.

Suggested citation: Mahler, D. & Tatulli, R. trackr: Tools to Evaluate
and Project Progress on Development Indicators. R package version 0.1.0.
Available at: github.com/RossanaTat/trackr
